---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Digital elevation model (100 m global)

[![lifecycle](https://img.shields.io/badge/Lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![DOI](https://zenodo.org/badge/430568284.svg)](https://zenodo.org/badge/latestdoi/430568284)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/README-", fig.align = "center")
```

This repository contains code to generate a global digital elevation model (based on Robinson *et al.* 2014). Specifically, elevation data are downloaded from the [EarthEnv project](https://www.earthenv.org/DEM.html). Next, the data are combined into a single raster (GeoTIFF) file containing the elevation data (90 m resolution; [EPSG:4326 coordinate reference system](https://epsg.io/4326)). Finally, the data are projected to the World Behrmann coordinate reference system using (100 m resolution; [ESRI:4326 coordinate reference system](https://epsg.io/54017)). A processed version of the data can be downloaded from the [Zenodo Digital Repository](TODO). **All credit should go to Robinson *et al.* (2014), please cite their work if you use this dataset.**

```{r "map", echo = FALSE, message = FALSE, warning = FALSE, fig.height = 2.5, fig.width = 7.0, dpi = 250}
# load packages
library(terra)
library(ggplot2)
library(viridis)

# load raster
x <- terra::rast("results/dem-100m-esri54017.tif")

# extract data to plot
plot_raster <- terra::spatSample(
  x, size = 400000, method = "regular", as.raster = TRUE
)
plot_data <- as.data.frame(plot_raster, xy = TRUE)
names(plot_data)[3] <- "value"
plot_data$title <- "Global elevation (100 m)"
plot_data$subtitle <- "dem-100m-esri54017.tif"
# make plot
p <-
  ggplot() +
  geom_tile(
    data = plot_data,
    mapping = aes(x = x, y = y, fill = value),
    height = yres(plot_raster),
    width = xres(plot_raster),
  ) +
  scale_fill_viridis(name = "Elevation (m)") +
  coord_fixed(expand = FALSE) +
  facet_wrap(~ title + subtitle) +
  theme(
    legend.pos = "right",
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.line = element_blank(),
    axis.ticks.length = unit(0, "null"),
    axis.ticks.margin = unit(0, "null"),
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid = element_blank(),
    legend.text = element_text(size = 8),
    legend.margin = unit(0, "null"),
    plot.margin = unit(c(0.001, 0.001, 0.001, 0.01), "pt"),
    strip.background = ggplot2::element_rect(color = "black", fill = "black"),
    strip.text = ggplot2::element_text(color = "white")
  )

# render plot
print(p)
```

## Metadata

The raster (GeoTIFF) file is `dem-100m-esri54017.tif` (see below for details).

```{r, echo = FALSE, comment = NA}
withr::with_dir("results", {
  cat(
    system("gdalinfo dem-100m-esri54017.tif", intern = TRUE),
    sep = "\n"
  )
})
```

## Usage

You can run the code to produce the dataset using the system command `make clean all`. After processing has successfully completed, the code will produce a raster (GeoTIFF) file named `dem-100m-esri54017.tif` in the `results` directory.

## Software requirements

Operating system
  * Ubuntu (21.10) (300 Gb disk space available)

Software
  * GNU make (version 4.1)
  * GDAL (version 3.2.2)
  * R (version 4.1.1)

System packages
  * gdal-bin
  * libudunits2-dev
  * libgdal-dev
  * libgeos-dev
  * libproj-dev
  * libxml2-dev
  * libssl-dev

## Citation

This dataset was derived from Robinson *et al.* (2014). If you use this dataset, you should cite their work - they deserve all the credit!

> Robinson N, Regetz J, & Guralnick RP (2014) EarthEnv-DEM90: A nearly-global, void-free, multi-scale smoothed, 90m digital elevation model from fused ASTER and SRTM data. *ISPRS Journal of Photogrammetry and Remote Sensing*, **87**: 57--67.
