---
output:
  rmarkdown::github_document:
    html_preview: no
---

<!--- README.md is generated from README.Rmd. Please edit that file -->

# Digital elevation model (100 m global)

[![lifecycle](https://img.shields.io/badge/Lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)

```{r, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/README-", fig.align = "center")
```

```{r, include = FALSE}
h = 3.5
w = 3.5
ow = "400"
```

This repository contains code to generate a global digital elevation model (based on Robinson *et al.* 2014). Specifically, elevation data are downloaded from the [EarthEnv project](https://www.earthenv.org/DEM.html). Next, the data are combined into a single raster (GeoTIFF) file containing the elevation data (90 m resolution; [EPSG:4326 coordinate reference system](https://epsg.io/4326)). Finally, the data are projected to the World Behrmann coordinate reference system using (100 m resolution; [ESRI:4326 coordinate reference system](https://epsg.io/54017)). Pre-processed versions of the data can be downloaded from a [Zenodo Digital Repository](TODO). **All credit should go to Robinson *et al.* 2014, please cite their work if you use this dataset.**

```{r}
library(terra)
r <- terra::rast("results/dem-100m-esri54017.tif")
plot(r, main = "Global elevation (100 m)")
```

# Usage

After cloning this repository, you can rerun the entire analysis on your own computer using the system command `make clean all`. Please see below for software requirements.

# Software requirements

* Operating system
  Ubuntu (21.1) (200 Gb disk space)
* Software
  GNU make (version 4.1)
  GDAL (version 3.2.2)
  R (version 4.1.1)
* System packages
  gdal-bin
  libudunits2-dev
  libgdal-dev
  libgeos-dev
  libproj-dev
  libxml2-dev
  libssl-dev

# References

Robinson N, Regetz J, & Guralnick RP (2014) EarthEnv-DEM90: A nearly-global, void-free, multi-scale smoothed, 90m digital elevation model from fused ASTER and SRTM data. *ISPRS Journal of Photogrammetry and Remote Sensing*, 87: 57--67.
